Analysis for Biracial Grade Retention Project

### 2022-09-03 analysis/analysis.Rmd
Specify weights manually in marg for counterfactual models
This is necessary because otherwise it tries to use the full weights which don't work for the new dataset

```{r test-marg, eval=FALSE}
# This code chunk just tests to make sure I get the same value by hand as I get 
# when I do it through the marg command.
temp <- subset(cf_df, race=="Black/White")
lor <- predict(model_full, newdata=temp)
mean(exp(lor)/(1+exp(lor)))
marg(model_full, "year", data=temp)
# It works!
```

@@

```{r create-cf-marg}
coefs_cf <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df, race==x)
  means_margin <- marg(model_full, "year", data=temp)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows()

@@

```{r counterfactual-ses, warning=FALSE, fig.height=12, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were resources (household income, parental education, home ownership, parent and child foreign-born status, parent and child English proficiency, and whether parents are married). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
ggplot(coef_table_cf, aes(x=term, y=estimate, color=multiracial,
                          ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table_cf, multiracial))+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")

@@

```{r cf-ses-isolate, warning=FALSE, fig.height=12, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were SES resources (household income, parental education, home ownership, and whether parents are married). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
cf_df_ses <- cf_df
cf_df_ses$foreign_born <- FALSE
cf_df_ses$foreign_born_father <- FALSE
cf_df_ses$foreign_born_mother <- FALSE
cf_df_ses$spk_eng <- "Yes"
cf_df_ses$spk_eng_father <- "Yes"
cf_df_ses$spk_eng_mother <- "Yes"

@@

```{r cf-culture-isolate, warning=FALSE, fig.height=12, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were cultural resources (foreign-born status and English proficiency). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
cf_df_culture <- cf_df
cf_df_culture$family_income <- mean(cf_df_culture$family_income)
cf_df_culture$own_home <- TRUE
cf_df_culture$degree_father <- "BA"
cf_df_culture$degree_mother <- "BA"
cf_df_culture$parents_married <- TRUE

>>>>

```{r test-marg, eval=FALSE}
# This code chunk just tests to make sure I get the same value by hand as I get 
# when I do it through the marg command.
temp <- subset(cf_df, race=="Black/White")
lor <- predict(model_full, newdata=temp)
weighted.mean(exp(lor)/(1+exp(lor)), w=temp$perwt)
marg(model_full, "year", data=temp, weights=temp$perwt)
# It works
```

@@

```{r create-cf-marg}
coefs_cf <- lapply(unique(acs$race), function(x) {
  temp <- subset(cf_df, race==x)
  means_margin <- marg(model_full, "year", data=temp, weights=temp$perwt)[[1]]
  return(tibble(term=x, estimate=means_margin$Margin, 
                se=means_margin$Standard.Error))
}) %>%
  bind_rows()

@@

```{r counterfactual-ses, warning=FALSE, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were resources (household income, parental education, home ownership, parent and child foreign-born status, parent and child English proficiency, and whether parents are married). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
ggplot(coef_table_cf, aes(x=term, y=estimate, color=multiracial,
                          ymin=estimate-1.386*se, ymax=estimate+1.386*se))+
  geom_rect(aes(xmax=Inf, xmin=-Inf, ymin=estimate_mid-1.386*se_mid,
                ymax=estimate_mid+1.386*se_mid, fill=multiracial), 
            color=NA, alpha=0.2)+
  geom_linerange(data=subset(coef_table_cf, multiracial))+
  geom_point()+
  coord_flip()+
  facet_wrap(~mrace, scales="free_y")+
  theme_bw()+
  theme(legend.position = "none", panel.grid = element_blank())+
  scale_color_manual(values=c("grey40","red"))+
  scale_fill_manual(values=c("grey40","red"))+
  scale_linetype_manual(values=c(3,1))+
  scale_y_continuous(labels = scales::percent)+
  labs(x=NULL,
       y="average adjusted predicted probability of being behind expected grade")

@@

```{r cf-ses-isolate, warning=FALSE, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were SES resources (household income, parental education, home ownership, and whether parents are married). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
cf_df_ses <- cf_df
cf_df_ses$foreign_born <- FALSE
cf_df_ses$foreign_born_father <- FALSE
cf_df_ses$foreign_born_mother <- FALSE
cf_df_ses$spk_eng <- "Yes"
cf_df_ses$spk_eng_father <- "Yes"
cf_df_ses$spk_eng_mother <- "Yes"

@@

```{r cf-culture-isolate, warning=FALSE, fig.cap="Counterfactual probability of being behind grade for biracial respondents in comparison to their monoracial comparison groups if the only differences between groups were cultural resources (foreign-born status and English proficiency). 83.4% confidence bands shown around each estimate. Non-overlap in confidence bands roughly indicates statistically significant difference at the 5% level."}
cf_df_culture <- cf_df
cf_df_culture$family_income <- mean(cf_df_culture$family_income)
cf_df_culture$own_home <- TRUE
cf_df_culture$degree_father <- "BA"
cf_df_culture$degree_mother <- "BA"
cf_df_culture$parents_married <- TRUE
